Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer

### 2020-04-26 manuscript/manuscript.tex
addessed TODOs regarding classification performance

A desirable solution offers good predictive power on unseen \emph{test} samples. We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy. We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control). The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier. This penalty was set by cross-validation, choosing the value that minimized misclassification error. We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs. When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes. Then we evaluated the sensitivity and the specificity on the \emph{test set}. The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution. To obtain a baseline, we also trained the classifier on all the SNPs. We do not expect a linear model on selected SNPs to be able to separate cases from controls well. Indeed, [TODO : can you add info about heritability of non-BRCA familial breast cancer, explain that only part of this heritability is going to be genetic, and conclude that we do not expect to look at a woman's SNPs and tell whether she has cancer or not?]. However, classification accuracy may be one additional informative criterion on which to evaluate solutions.

@@

We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}). Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime. The solution size varies greatly between the different methods (Figure~\ref{fig:benchmark}A). Heinz produced the smallest solutions, with an average of 182 selected SNPs. The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs). To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B). The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range. Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error. \cazcom{}{TODO : Add a sentence to the effect of ``Although these classification performance may seem low, remember we are not really hoping to separate cases from controls well from SNP data''?}

>>>>

A desirable solution offers good predictive power on unseen \emph{test} samples. We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy. We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control). The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier. This penalty was set by cross-validation, choosing the value that minimized misclassification error. We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs. When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes. Then we evaluated the sensitivity and the specificity on the \emph{test set}. The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution. To obtain a baseline, we also trained the classifier on all the SNPs. We do not expect a linear model on selected SNPs to be able to separate cases from controls well. Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}. However, classification accuracy may be one additional informative criterion on which to evaluate solutions.

@@

We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}). Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime. The solution size varies greatly between the different methods (Figure~\ref{fig:benchmark}A). Heinz produced the smallest solutions, with an average of 182 selected SNPs. The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs). To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B). The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range. Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error. It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
